library(knitr)
library(ggplot2)
library(grid)
library(gridExtra)
# Setting the home directory and reading the activity csv file  

```{r}  
setwd("/Users/Sandman/MOOC/Reproducible Research/")  
rawdata <- read.csv("activity.csv",header=TRUE,sep=",")  
```

# Cleaning up data and creating histogram, mean and median  
```{r,echo=TRUE}  
cleandata <- rawdata[!is.na(rawdata$steps),]  
sum_clean_data <- aggregate(cleandata$steps,by=list(date=cleandata$date),FUN=sum)  
names(sum_clean_data) = c("Date","Steps")  
hist(sum_clean_data$Steps)  
utils::str(hist(sum_clean_data$Steps, col = "gray", labels = TRUE))  
mean_clean_data <- mean(sum_clean_data$Steps)  
median_clean_data <- median(sum_clean_data$Steps)
mean_clean_data
median_clean_data
```

# Calculating Daily Active Pattern  
```{r,echo=TRUE}  
avg_clean_data <- aggregate(cleandata$steps,by=list(date=cleandata$interval),FUN=mean)  
names(avg_clean_data) = c("Interval","Average_Steps")  
plot(avg_clean_data$Interval,avg_clean_data$Average_Steps,type="l")  
avg_clean_data[avg_clean_data$Average_Steps==max(avg_clean_data$Average_Steps),]   
```



# Inputting missing values
    
## Counting total no of records with NA  
```{r,echo=TRUE}  
Total_NAs <- nrow(rawdata[is.na(rawdata$steps),])  
Total_NAs  
```  

## Filling in missing values with mean values
 
```{r,echo=TRUE}
names(rawdata) <- c("Steps","date","Interval")
head(rawdata)
rawdata_mod <- merge(rawdata,avg_clean_data,all.x=TRUE)
rawdata_mod_na <- rawdata_mod[is.na(rawdata_mod$Steps),]
names(rawdata) <- c("Steps","date","Interval")
rawdata_mod <- merge(rawdata,avg_clean_data,all.x=TRUE)
rawdata_mod_na <- rawdata_mod[is.na(rawdata_mod$Steps),c("Interval","date","Average_Steps")]
names(rawdata_mod_na) <- c("Interval","date","Steps")
rawdata_mod_clean <- rawdata_mod[!is.na(rawdata_mod$Steps),c("Interval","date","Steps")]
```

## Dataset with filled in values are stored in a data frame
```{r,echo=TRUE}
rawdata_mod_final <- rbind(rawdata_mod_clean,rawdata_mod_na)
names(rawdata_mod_final) <- c("Interval","date","steps")
```


## Show median and mean values with filled in data
```{r,echo=TRUE}
sum_mod_data <- aggregate(rawdata_mod_final$steps,by=list(date=rawdata_mod_final$date),FUN=sum)
names(sum_mod_data) = c("Date","Steps")

hist(sum_mod_data$Steps)
utils::str(hist(sum_mod_data$Steps, col = "gray", labels = TRUE))

mean_mod_data <- mean(sum_mod_data$Steps)
median_mod_data <- median(sum_mod_data$Steps)
```



# Are there differences in activity patterns between weekdays and weekends?
## Creating factor variable with 2 values - Weekdays and Weekends

```{r,echo=TRUE}
filleddata <- cbind(rawdata_mod_final,dayofweek = ifelse(weekdays(as.Date(rawdata_mod_final$date),abbreviate=FALSE) %in% c("Saturday","Sunday"),"Weekend","Weekday"))

filleddata_wd <- filleddata[filleddata$dayofweek=="Weekend",]
filleddata_we <- filleddata[filleddata$dayofweek=="Weekday",]

weekday_filled_data <- aggregate(filleddata_wd$steps,by=list(Interval=filleddata_wd$Interval),FUN=mean)
plot(weekday_filled_data$Interval,weekday_filled_data$x,type="l")

weekend_filled_data <- aggregate(filleddata_we$steps,by=list(Interval=filleddata_we$Interval),FUN=mean)
plot(weekend_filled_data$Interval,weekend_filled_data$x,type="l")
```

## Create a panel plot with average number of steps taken across weekend days and weekdays
```{r,echo=TRUE}
a <- ggplot() +
  geom_line(data=weekday_filled_data, aes(x = Interval,y = x),colour = "Red")
b<- ggplot() +
  geom_line(data=weekend_filled_data, aes(x = Interval,y = x),colour = "Green")

grid.arrange(a,b,ncol=1)
ggplot(data=filleddata,aes(x= Interval,y= steps))+geom_line() + facet_grid(dayofweek ~ .)
```

